\textbf{On the Common Logical Structure of Classical and Quantum Mechanics}

### 2021-10-10 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: adding comments after review

The above proposition contains our first crucial conclusion. As we stated in Section \ref{QL}, in quantum logic the propositions correspond to closed subspaces, which are closed sets and, thus, are Borel sets. All the propositions of quantum logic are consequently readily found in the Borel algebra of the Hilbert space. The Borel algebra, in turn, is a Boolean algebra that follows classical logic, and contains all the statements of quantum logic. This construction also applies identically for the classical cases, where we will also be able to talk about the set of all closed subspaces.

@@

Since $\langle \psi , A \psi \rangle$ expresses the expected value for operator $A$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators. However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limited. Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity. Renormalization would lead to the set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $\mathcal{H} \setminus \{ 0 \}$.} Consider, for example, ``the expectation of $z$-spin is zero''. This, by Proposition 2, would correspond to a Borel set $V \subseteq \mathcal{H}$. This set is not the full set since the state with $z$-spin up is not in the set. However, $x$-spin up and $x$-spin down will be in the set, and their span is the full space. Therefore $V$ is not a closed subspace. Our proposition, despite being physically interesting, does not correspond to a closed subspace.

The idea is that closed linear subspaces correspond to single-shot measurements. While these form an interesting subset of possible measurements, they are fairly limited in what they can tell us about the incoming system. For example, measuring $z$-spin up tells us very little about spin \textbf{before} the measurement: only that it couldn't be $z$-spin down. To make any inference on the incoming state we need to gather statistics. In fact, no cross section measurement, our main tool in exploring the standard model, is single-shot. Being that quantum mechanics is a statistical theory, most physically interesting statements are statistical in nature, and that is why we need to include the Borel sets in our lattice of statements.

>>>>

The above proposition contains our first crucial conclusion. As we stated in Section \ref{QL}, in quantum logic the propositions correspond to closed subspaces, which are closed sets and, thus, are Borel sets. All the propositions of quantum logic are consequently readily found in the Borel algebra of the Hilbert space. The Borel algebra, in turn, is a Boolean algebra that follows classical logic, and contains all the statements of quantum logic.\footnote{Obviously, the inclusion map from the lattice of closed subspaces is not an order isomorphism as the join and complement are not preserved. However, these are still expressible as subspace closure and subspace complement.} This construction also applies identically for the classical cases, where we will also be able to talk about the set of all closed subspaces.

@@

Since $\langle \psi , A \psi \rangle$ expresses the expected value for operator $A$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators. However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limited. Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity. Renormalization would lead to the set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $\mathcal{H} \setminus \{ 0 \}$.} Consider, for example, ``the expectation of $z$-spin is zero''. This would correspond to the set $V \subseteq \mathcal{H}$ that includes all states, and only the states, for which the expection of $S_z$ is zero. By Proposition 2, $V$ is a Borel set. In fact, the set $U = {0}$ is a Borel set of $\mathbb{R}$ and therefore $V = F_{S_z}^{-1}(U)$ is a Borel set of $\mathcal{H}$. However, $V$ is not a closed subspace. The set $V$ is not the full set $\mathcal{H}$ since, for example, the state with $z$-spin up is not in the set. However, $x$-spin up and $x$-spin down will be in the set $V$, and their span is the full space $\mathcal{H}$. Therefore $V$ is not a closed subspace. Our proposition, despite being physically interesting, does not correspond to a closed subspace but does correspond to a Borel set.

The idea is that closed linear subspaces correspond to single-shot measurements. While these form an interesting subset of possible measurements, they are fairly limited in what they can tell us about the incoming system. For example, measuring $z$-spin up tells us very little about spin \textbf{before} the measurement: only that it couldn't be $z$-spin down. To make any inference on the incoming state we need to gather statistics. In fact, no cross section measurement, our main tool in exploring the standard model, is single-shot. Being that quantum mechanics is a statistical theory, most physically interesting statements are statistical in nature, and that is why we need to include the Borel sets in our lattice of statements.\footnote{Since probabilities are assigned to Borel sets and only Borel sets, the Borel algebra will contain all physically interesting statements. Further expansions (e.g. to the power set) would only include unphysical statements.}
